Cover letter for the original IJRR submission

### 2017-09-29 reviews_tro/reviewer_response.tex

\documentclass[a4paper]{article}
 \usepackage{fullpage}
\usepackage{amsfonts} % pour les lettres maths creuses \mathbb
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[frenchb]{babel}
\usepackage{aeguill}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{color}
\usepackage{listings}
\usepackage{pdfpages}
% Another method to track changes
% Examples of usage:
% - This is \added[id=per,remark={we need this}]{new} text.
% - This is \deleted[id=per,remark=obsolete]{unnecessary}text.
% - This is \replaced[id=per]{nice}{bad} text.
% To print the list of change use: \listofchanges
%~ \usepackage{changes}	% use this for the working version
%\usepackage[final]{changes} % use this for the final version
%~ \definechangesauthor[name={Andrea Del Prete}, color=orange]{adp}
%~ \newcommand{\deladp}[1]{\deleted[id=adp]{#1}}
%~ \newcommand{\addadp}[1]{\added[id=adp]{#1}}
%~ \newcommand{\repadp}[2]{\replaced[id=adp]{#1}{#2}}
%~ \definechangesauthor[name={Steve Tonneau}, color=blue]{st}
%~ \newcommand{\delst}[1]{\deleted[id=st]{#1}}
%~ \newcommand{\addst}[1]{\added[id=st]{#1}}
%~ \newcommand{\repst}[2]{\replaced[id=st]{#1}{#2}}
%~ \newcommand{\gls}[1]{\textit{#1}}
%~ \newcommand{\glslink}[2]{{#2}}
\newcommand{\done}[0]{\textcolor{green}{DONE}}
\newcommand{\ndone}[0]{\textcolor{red}{TODO}}

\newcommand\dx{\dot{x}}
\newcommand\dy{\dot{y}}
\newcommand\ddx{\ddot{x}}
\newcommand\ddy{\ddot{y}}
\def\real{{\mathbb R} }
\newcommand\quot[1]{\begin{quote} \underline{quote}: \textbf{#1}\end{quote}}
\newcommand\as[1]{\begin{quote} \underline{answer}: {#1}\end{quote} \leavevmode \\ }
\newcommand\qt[1]{\begin{quote} \underline{added text in paper}: \textit{#1}\end{quote} \leavevmode \\ }
\frenchbsetup%
{%
StandardItemLabels=true,%
ItemLabels=\ding{43},%
}%
\DeclareUnicodeCharacter{00A0}{ }
\author {}
\title {Cover letter for the resubmission of our paper ``An efficient acyclic contact planner for multiped robots''}
\date {}
\begin{document}
\maketitle

@@

\section{Answers to reviewer 1}

\quot {As for novel contributions, this did not seem to be addressed/explained
until page 3 (in the last paragraph of Sec. I). Much of the methodology
has already been presented in earlier work, and the particular aspects
that set THIS paper apart seem (as claimed by the authors) to be: 
}

\as{Thanks for the comment. Indeed, all reviewers agreed that the contribution with respect to previous work was not clearly stated. We updated the paper to reflect the following. Firstly, our planner is almost identical to the one in [20], with important additions that are the pseudo code for the algorithm (the algorithm is not given in [20]), as well as the introduction of a more efficient criterion
for checking static equilibrium (see next question). The optimal control is no longer a claimed contribution of the paper, even though it differs from [1] in its formulation. The main
novelty of the paper resides in the experimental setup and analysis, as well as the empirical proof that the planner works with real robots that are more constrained kinematically than the virtual avatars from [20]. We believe that this paper is thus in accordance with the guidelines of T-RO regarding the extension of a conference paper.}
\qt{The present paper is an extension of our conference paper[20]. As such, the motion planning algorithm presented here is the same as the one presented in the
previous work. Section IV and V of the present paper are thus paraphrasing the original paper, with the exception of Section V.B.2, where we additionally provide the pseudo-code of the algorithm absent from the original paper. We also improve the computational performance of the method by replacing the static equilibrium test with a more efficient test based on an new LP formulation presented in Section 6. \\
Compared with the original paper, we also release our source code, and provide the complete procedure to use our planner for any legged robot. We thus provide
all the technical information required to use the planner. \\
To validate our approach, in this paper we had to implement or integrate other tools based on previous publications from our team. We implemented a solution to $\mathcal{P}_3$ based on a reformulation of our previous work[1]. The implementation of this solution is also provided as an open-source project, which can be integrated directly in our framework. Our solution to $\mathcal{P}_3$ then allows us to verify that the synthesized motions are physically consistent, using our implementation of the state of the art simulator proposed in \textcolor{red}{TODO}.
We do not consider these additions as contributions per se, but describing them here is required to understand the experimental contribution of this paper.
Indeed, thanks to these tools we have been able to design a complete pipeline for synthesizing multi contact motions from scratch.
The novelty of the present paper lies in the rigorous experimental setup we propose to demonstrate its interest.
As opposed to our previous paper, we demonstrate the algorithms with actual robotic models. These models are much more constrained kinematically than the virtual avatars used before,
thus making the planning problem even harder. Through an extensive empirical analysis, we demonstrate that the method remains interesting in this context, qualitatively and efficiently.}
\done

%~ \as{We agree that our original submission is not clear on the novelty between this paper and previous contributions.
%~ This paper must be understood as an extension of the ISRR paper [20], while the other cited contributions from our team 
%~ are exploited to validate our approach in a complete pipeline.
%~ In this extension, the algorithm used to compute a sequence of contacts is the same as in [20] (although this
%~ algorithm is not explicitely given in [20]), with the introduction of a fastest test for verifying the robust quasi static equilibrium of a contact posture.
%~ We reformulated the end of Section I.B to state clearly that the novelty of this paper essentially lies in the experimental setup and validation we achieved with real robot models, in particular
%~ in the empirical justification of the method, and its validation in a dynamic simulation. This effort led to the proposal of the first open source, complete framework for multi contact motion synthesis, since 
%~ the validation required to write a solver for the problem P3. }\done

@@

\quot{I do, however,
believe that the fact that the technical approaches are rephrasings of
prior work should be made more clear (or that sections which can be
replaced by citations of [20] and [1] be removed).}
\as{Thanks for the advice. This suggestion was in contradiction with the other reviewers, who thought that the paper should be self-contained.
The associate editor decided that the paper must be self contained, so we maintained the indicated sections, but made clearer the fact that they are rephrasing of previous work. }\ndone

\quot{*I think I understand what the authors mean by solving at
"interactive" rates, but this is worth defining. In my mind, it means
something like "before the human operator gets bored", or roughly > 0.5
Hz. But since the authors make the claim that this is the "first
interactive implementation of a contact planner", then that statement
should be made quantitative and verifiable}
\as{We agree with the relevant comment, and updated the paper accordingly.}\ndone

\quot{* In Section I.A. the authors refer to Bretl's work on multi-contact
planning as the inspiration for the P1,P2,P3 decomposition of the
problem. But it is not, as the authors state, always necessary to
decompose motion planning into choosing a series of quasi-static
configurations. There are a variety of feasible motions with no static
equilibrium configurations, so it is worth mentioning that this
decomposition only solves a subset of motion planning problem.}
\as{We apologize if this was not made clear enough in our first submission. In the conclusion, we indicate that the planner is limited to specific scenarios (where quasi-static motion
is always an option). We updated the state of the art to indicate that the decomposition in general is restricitve.}\ndone

\quot{* Also in Sect. I.A, the authors state that "optimization-based
approaches only converge locally". This is not necessarily true, as
convex or mixed-integer convex optimizations can (sometimes) be solved
to a verifiable global optimum. I would suggest removing this sentence
entirely.}
\as{Thanks for the suggestion. Our phrasing was indeed misleading and inaccurate. Our intention was to explain that the geometry of the scene makes the problem highly non-convex, and
that approaches such as the CIO, if not initialized with an appropriate initial guess, might never get out of the local minima. We followed the reviewer's advice and simply removed the sentence. }\ndone

### 2017-09-29 reviews_tro/reviewer_response.tex

\documentclass[a4paper]{article}
 \usepackage{fullpage}
\usepackage{amsfonts} % pour les lettres maths creuses \mathbb
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[frenchb]{babel}
\usepackage{aeguill}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{color}
\usepackage{listings}
\usepackage{pdfpages}
% Another method to track changes
% Examples of usage:
% - This is \added[id=per,remark={we need this}]{new} text.
% - This is \deleted[id=per,remark=obsolete]{unnecessary}text.
% - This is \replaced[id=per]{nice}{bad} text.
% To print the list of change use: \listofchanges
%~ \usepackage{changes}	% use this for the working version
%\usepackage[final]{changes} % use this for the final version
%~ \definechangesauthor[name={Andrea Del Prete}, color=orange]{adp}
%~ \newcommand{\deladp}[1]{\deleted[id=adp]{#1}}
%~ \newcommand{\addadp}[1]{\added[id=adp]{#1}}
%~ \newcommand{\repadp}[2]{\replaced[id=adp]{#1}{#2}}
%~ \definechangesauthor[name={Steve Tonneau}, color=blue]{st}
%~ \newcommand{\delst}[1]{\deleted[id=st]{#1}}
%~ \newcommand{\addst}[1]{\added[id=st]{#1}}
%~ \newcommand{\repst}[2]{\replaced[id=st]{#1}{#2}}
\newcommand{\gls}[1]{\textit{#1}}
\newcommand{\glslink}[2]{{#2}}
\newcommand{\done}[0]{\textcolor{green}{DONE}}
\newcommand{\ndone}[0]{\textcolor{red}{TODO}}

\newcommand\dx{\dot{x}}
\newcommand\dy{\dot{y}}
\newcommand\ddx{\ddot{x}}
\newcommand\ddy{\ddot{y}}
\def\real{{\mathbb R} }
\newcommand\quot[1]{\begin{quote} \underline{quote}: \textbf{#1}\end{quote}}
\newcommand\as[1]{\begin{quote} \underline{answer}: {#1}\end{quote} }
\newcommand\qt[1]{\begin{quote} \underline{added text in paper}: \textit{#1}\end{quote} \leavevmode \\ }
\newcommand\jp{ \leavevmode \\}
\frenchbsetup%
{%
StandardItemLabels=true,%
ItemLabels=\ding{43},%
}%
\DeclareUnicodeCharacter{00A0}{ }
\author {}
\title {Cover letter for the resubmission of our paper ``An efficient acyclic contact planner for multiped robots''}
\date {}
\begin{document}
\maketitle

@@

%~ \as{We agree that our original submission is not clear on the novelty between this paper and previous contributions.
%~ This paper must be understood as an extension of the ISRR paper [20], while the other cited contributions from our team 
%~ are exploited to validate our approach in a complete pipeline.
%~ In this extension, the algorithm used to compute a sequence of contacts is the same as in [20] (although this
%~ algorithm is not explicitely given in [20]), with the introduction of a fastest test for verifying the robust quasi static equilibrium of a contact posture.
%~ We reformulated the end of Section I.B to state clearly that the novelty of this paper essentially lies in the experimental setup and validation we achieved with real robot models, in particular
%~ in the empirical justification of the method, and its validation in a dynamic simulation. This effort led to the proposal of the first open source, complete framework for multi contact motion synthesis, since 
%~ the validation required to write a solver for the problem P3. }\done

\quot { (1) A criterion for robust static equilibrium, but it isn't
described until page 12, in Appendix B, Section B. [\dots]  It would really help to describe this MUCH earlier in
the paper (than page 13), if it is a significant novelty of your paper.}

\as{We agree with the remark. To better distinguish between what is novel and not, we modified the paper to move the content of Appendix B.B (now Section VI), as well as D  (the pseudo code of our algorithm, now Section V.B.2) and E (reference to our source code, now Section VII) to the text itself. This results in the criterion being presented much earlier.}\done \jp

@@

\quot{It is a pretty intuitive/typical check for quantifying a
margin for friction conesque violations (using an approximation for
the friction cone, with x and y direction forces considered
independently). }

\as{Thanks for the comment. To our knowledge, what is common in the literature is to reduce the friction coefficient to obtain more conservative bounds on the admissible forces, and thus a ``safety margin'' to the real cone boundary. An issue with this formulation is that the resulting margin is really small near the origin of the cone, and really important as we go further away from it. With our formulation, the obtained margin is constant everywhere on the cone, and thus guaranteed to be greater than the value $b_0$ we compute. We believe this formulation is new, but would be happy to cite any relevant reference if we happen to be wrong. This formulation proves useful to us because the margin $b_0$ allows to choose the more robust candidate for static equilibrium.}\done

\quot{
Provides complete pseudocode.  This is great! However,it's
not fully clear how much of this is novel vs just re-presenting the
same algorithms previously publilshed.	If only code in App. B and C
are new, that should be clear. Conversely, if most of this presentation
was NOT outlined in such detail before, that should also be clear (to
the authors' credit). Either way, a brief clarification in the intro is
recommended.}

\as{We agree with the comment. Indeed, the pseudocode was not present in [20]. We included it in the core of the paper and clarified that it is novel at the end
of Section I.B.}
\qt{Section IV and V of the present paper are thus paraphrasing the original paper, with the exception of Section V.B.2, where we additionally provide the pseudo-code of the algorithm absent from the original paper.}
\done

@@

\quot{(3)	The contribution to provide a solution to P3 doesn't happen
until page 13 (Appendix C) and as with contribution (1), it needs a
better, intuitive summary (briefly) when first mentioned, as well as a
roadmap for the reader as to where it will be described. It is also
ambiguous in the paper whether to what extent (if any) this solution is
novel, versus a presentation of work from [1]. "Provide a solution"
seemed to imply novelty, in the context of this paragraph -- yet later,
the authors (near the end of page 2) state they use a "state-of-the-art
solution to P3", which implies the P3 solution is not a claim of
novelty.}

\as{The main contribution of the paper is the proposal and validation of an algorithm for P1 and P2. 
In the new version of the paper, we do not consider that our solution to P3
is a new contribution, although our formulation differs from [1] (our control variable is the acceleration
of the center of mass, where in [1] the control variable is the contact forces). We provide it here for completeness, because it corresponds to our implementation, and 
was required to validate our approach. 
We rephrased the text to clearly indicate that the algorithm is not a contribution. We made the choice not to give an intuition nor describe the solver in the core
of the paper because it is not required to understand the contributions to P1 and P2.}
\qt{ We implemented a solution to $\mathcal{P}_3$ based on a reformulation of our previous work[1]. The implementation of this solution is also provided as an open-source project, which can be integrated directly in our framework. Our solution to $\mathcal{P}_3$ then allows us to verify that the synthesized motions are physically consistent, using our implementation of the state of the art simulator proposed in \textcolor{red}{TODO}.
We do not consider these additions as contributions per se, but describing them here is required to understand the experimental contribution of this paper.}\done

@@

\quot{[\dots] the authors need to better
distinguish what the novel contributions are, and to highlight where
these contributions will be presented in the paper.  Also, many
important details are in the appendices. The paper would be improved
with more intuitive comments/summaries on these details, rather than
just writing (for example), we use a bunch of heuristics, and they are
described in Appendix X.}

\as{We agree. The novel details of the paper have been moved from the appendix to the core of the paper, and are explicitely detailed at the end of Section I.B.} \done

\quot{
Also, where you describe "sensitivity" vs "specificity", it would be
useful to give a more intuitive explanation, example [\dots]}

\as{We updated Section VIII.B.1) to illustrate the meaning of those terms.}
\qt{We then computed to sensitivity and specificity of the reachability condition.  In this context, the sensitivity refers to the percentage of configurations in $C_{Reach}^0$, effectively belonging to \gls{$C_{Contact}^0$}. If a sampled configuration is in $C_{Reach}^0$, but our method is unable to generate a contact configuration from it, as a result the sensitivity
decreases. The sensitivity thus illustrates the confidence we have that any configuration in $C_{Reach}^0$ will effectively lead to a contact configuration.
Conversely, the specificity refers to the percentage of configurations \textbf{not} in $C_{Reach}^0$, effectively \textbf{not} belonging to \gls{$C_{Contact}^0$}.
If a sampled configuration is \textbf{not} in $C_{Reach}^0$, but our method is able to generate a contact configuration from it, as a result the specificity
decreases. The specificity thus illustrates the confidence we have that all configurations that allow contact creation belong to in $C_{Reach}^0$ (or informally, the confidence that we are not missing too many valid solutions).
We thus look for a compromise between sensitivity and specificity.} \done

### 2017-09-29 body/intro.tex
typo

\noindent \textbf{Comparison with our previous work:}
The present paper is an extension of our conference paper~\cite{tonneauisrr15}. As such, the motion planning algorithm presented here is the same as the one presented in the
previous work. Section~\ref{rbprm} and~\ref{sec:contact} of the present paper are thus paraphrasing the original paper, with the exception of Section~\ref{app:contact}, where we additionally provide the pseudo-code of the algorithm absent from the original paper. We also improve the computational performance of the method by replacing the static equilibrium test with a more efficient test based on an new LP formulation presented in Section~\ref{sec:Equil}..
Compared with the original paper, we also release our source code, and provide the complete procedure to use our planner for any legged robot. We thus provide
all the technical information required to use the planner.

>>>>

\noindent \textbf{Comparison with our previous work:}
The present paper is an extension of our conference paper~\cite{tonneauisrr15}. As such, the motion planning algorithm presented here is the same as the one presented in the
previous work. Section~\ref{rbprm} and~\ref{sec:contact} of the present paper are thus paraphrasing the original paper, with the exception of Section~\ref{app:contact}, where we additionally provide the pseudo-code of the algorithm absent from the original paper. We also improve the computational performance of the method by replacing the static equilibrium test with a more efficient test based on an new LP formulation presented in Section~\ref{sec:Equil}.
Compared with the original paper, we also release our source code, and provide the complete procedure to use our planner for any legged robot. We thus provide
all the technical information required to use the planner.
